Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-21 paper.tex
Add Jannis's acknowledgements
Save a line in abstract

\begin{abstract}
  Several proof assistants provide automation tactics based on tableau-style tree search, such as Isabelle's and Rocq's auto and Lean's Aesop.
  In this setting we consider \emph{forward rules}, which apply a given theorem, say, $A → B → C$, to any goal containing hypotheses $A$ and $B$, adding $C$ as a new hypothesis.
  When treated naively, such rules are tried on every goal encountered during the search, leading to repeated unifications of premises $A$ and $B$ with the hypotheses of each goal.
  We present an approach to forward rules that avoids some of this repeated work by taking advantage of similarities between successive goals.
  For each goal, we cache partial applications of forward rules in a custom data structure, the \emph{forward state}, which enables efficient updates.
  Our technique is compatible with any search strategy and most logics.
  It has been implemented in Aesop.

@@

>>>>

\begin{abstract}
  Several proof assistants provide automation tactics based on tableau-style tree search, such as Isabelle's and Rocq's auto and Lean's Aesop.
  In this setting we consider \emph{forward rules}, which apply a given theorem, say, $A → B → C$, to any goal containing hypotheses $A$ and $B$, adding $C$ as a new hypothesis.
  When treated naively, such rules are tried on every goal encountered during the search, leading to repeated unifications of premises $A$ and $B$ with the hypotheses of each goal.
  We present an approach to forward rules that avoids some of this repeated work by taking advantage of similarities between successive goals.
  For each goal, we cache partial applications of forward rules in a custom data structure that enables efficient updates.
  Our technique is compatible with any search strategy and most logics.
  It has been implemented in Aesop.

@@

\para{Acknowledgements.}
We thank Jasmin Blanchette and Massin Guerdi for their valuable comments on drafts of this paper.
\jcom{TODO Xavier funding.}
